---
title: "Homework-number-5"
author: "Nandini Sharma, Nitin Verma"
date: "November 18, 2016"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction 
When we initially discussed the central limit theorem and confidence intervals, we showed how we could use bootstrapping to estimate standard errors and confidence intervals around certain parameter values, like the mean. Using bootstrapping, we could also do the same for estimating standard errors and CIs around regression parameters, such as $\beta$ coefficients.

#Linear Regression{.tabset .tabset-fade .tabset-pills}

##Reading Data
```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/difiore/ADA2016/master/KamilarAndCooperData.csv")
d <- read.csv(f, header=TRUE, stringsAsFactors=TRUE)
head(d)
```
##Part 1 - Regression Model
**Using the "KamilarAndCooperData.csv" dataset, run a linear regression looking at log(HomeRange_km2) in relation to log(Body_mass_female_mean) and report your β coefficients (slope and intercept).**
```{r}
y <- log(d$HomeRange_km2)
x <- log(d$Body_mass_female_mean)

#Building the Model
m <- lm(data=d, y ~ x)
summary(m)

#Extracting Beta Coefficients
beta0 <- m$coefficients[1]
beta1 <- m$coefficients[2]

beta0
beta1
```

## Part 2 - BootStrapping {.tabset .tabset-fade .tabset-pills}
**Then, use bootstrapping to sample from your data 1000 times with replacement, each time fitting the same model and calculating the same coefficients. This generates a sampling distribution for each β coefficient.**
```{r}
#Defining Sample Size
n <- nrow(d)

beta0s <- NULL
beta1s <- NULL

#Bootstrapping for Beta Coefficients
for (i in 1:1000) {
  ds <- d[sample(c(1:n), n, replace=TRUE), ]
  ys <- log(ds$HomeRange_km2)
  xs <- log(ds$Body_mass_female_mean)
  ms <- lm(data=ds, ys ~ xs)
  
  beta0s[i] <- ms$coefficients[1]
  beta1s[i] <- ms$coefficients[2]
}
```

# Bootstrapping SE&CI {.tabset .tabset-fade .tabset-pills}
**Estimate the standard error for each of your β coefficients as the standard deviation of the sampling distribution from your bootstrap and determine the 95% CI for each of your β coefficients based on the appropriate quantiles from your sampling distribution.**
```{r}
#Standard Error from Bootstrap
se.beta0s <- sd(beta0s)
se.beta1s <- sd(beta1s)
se.beta0s
se.beta1s

#critval.beta0s <- qt(0.975, df= length(beta0s) - 1)
#critval.beta1s <- qt(0.975, df= length(beta1s) - 1)

#CI.beta0s = mean(beta0s) + c(-1, 1) * critval.beta0s * se.beta0s
#CI.beta1s = mean(beta1s) + c(-1, 1) * critval.beta1s * se.beta1s

#Calculating Confidence Intervals
CI.beta0s = quantile(beta0s, c(0.025, 0.975))
CI.beta1s = quantile(beta1s, c(0.025, 0.975))

CI.beta0s
CI.beta1s
```

## Comparing SE
**How does the former compare to the SE estimated from your entire dataset using the formula for standard error implemented in lm()?**

The SEs we calculated from bootstrap for `beta0` (`r se.beta0s`) and `beta1` (`r se.beta1s`) compare very well to the figures generated by `lm()` i.e. (`r summary(m)$coefficients[,"Std. Error"]`).

## Comparing CI
**How does the latter compare to the 95% CI estimated from your entire dataset?**

The CIs for the model `m` gotten via `lm()` can be calculated as follows:
```{r}
#CI.lm <- confint(m, m$coefficients[1], 0.95)
CI.lm <- confint(m, level = 0.95)
CI.lm
```
And these CIs compare very well with those obtained from bootstrapped sampling distributions i.e. beta0 **(`r CI.beta0s`)** and beta1 **(`r CI.beta1s`)**.

# Encapsulating Function
**Write a FUNCTION that takes as its arguments a dataframe, "d", a linear model, "m" (as a character string, e.g., "logHR~logBM"), a user-defined confidence interval level, "conf.level" (with default = 0.95), and a number of bootstrap replicates, "n" (with default = 1000). Your function should return a dataframe that includes: beta coefficient names; beta coefficients, standard errors, and upper and lower CI limits for the linear model based on your entire dataset; and mean beta coefficient estimates, SEs, and CI limits for those coefficients based on your bootstrap.**

```{r}
betaCoeffAnalysis = function(d1, m1, conf.level = 0.95, n = 1000){
  #Error handling for invalid dataframe/model
  if (nrow(d1) == 0 || is.null(m1)) {
    cat("ERROR!")
    return(NULL)
  }
  #Error handling for invalid confidence level
  if (conf.level > 1 || conf.level < 0) {
    cat("ERROR!")
    return(NULL)
  }
  #Extracting response and predictor variables from model string
  y1 <- strsplit(m1, "~")[[1]][[1]] #response variable
  x1 <- strsplit(m1, "~")[[1]][[2]] #predictor variable
  
  y <- d1[, c(y1)]
  x <- d1[, c(x1)]
  
  #Preparing result dataframe
  rdf <- data.frame(row.names = c("method", "beta", "est.value", "SE", "CI.lower", "CI.upper"))
  
  #Building regression model
  m <- lm(y ~ x)
  out <- summary(m)
  
  # Extracting coefficients and standard error from model
  b0.lm <- out$coefficients[1,1]
  se.b0.lm <- out$coefficients[1,2]
  b1.lm <- out$coefficients[2,1]
  se.b1.lm <- out$coefficients[2,2]
  
 # alpha <- conf.level + (1 - conf.level)/2
  
  # Extracting confidence interval from model
  CI.lm <- confint(m, level=conf.level)
  
  b0.lm.lwr <- CI.lm[1,1]
  b0.lm.upr <- CI.lm[1,2]
  b1.lm.lwr <- CI.lm[2,1]
  b1.lm.upr <- CI.lm[2,2]
  
  #Binding a record of model estimates for beta0 to the result dataframe rdf
  newrecord <- data.frame(method="lm", beta="beta0", est.value=b0.lm, SE=se.b0.lm, CI.lower=b0.lm.lwr, CI.upper=b0.lm.upr)
  rdf <- rbind(rdf, newrecord)
  
  #Binding a record of model estimates for beta1 to the result dataframe rdf
  newrecord <- data.frame(method="lm", beta="beta1", est.value=b1.lm, SE=se.b1.lm, CI.lower=b1.lm.lwr, CI.upper=b1.lm.upr)
  rdf <- rbind(rdf, newrecord)
  
  #Bootstrapping beta0 and beta1 estimates
  beta0s <- NULL
  beta1s <- NULL
  
  #Defining sample size for bootstrapping
  sz <- nrow(d1)
  for (i in 1:n) {
    
    ds <- d[sample(c(1:sz), sz, replace=TRUE), ]
    ys <- ds[, c(y1)]
    xs <- ds[, c(x1)]
    ms <- lm(ys ~ xs)
  
    beta0s[i] <- ms$coefficients[1]
    beta1s[i] <- ms$coefficients[2]
  }
  # Bootstrapped Mean and standard error for beta 0 and beta 1
  b0.bs <- mean(beta0s)
  b1.bs <- mean(beta1s)
  se.b0.bs <- sd(beta0s)
  se.b1.bs <- sd(beta1s)
  
#  critval.b0 <- qt(alpha, df= length(beta0s) - 1)
#  critval.b1 <- qt(alpha, df= length(beta1s) - 1)
  
#  CI.b0 = b0.bs + c(-1, 1) * critval.b0 * se.b0.bs
#  CI.b1 = b1.bs + c(-1, 1) * critval.b1 * se.b1.bs
  
  #Bootrapping confidence interval
  
  CI.b0 <- quantile(beta0s, level = conf.level)
  CI.b1 <- quantile(beta1s, level = conf.level )
  
  b0.bs.lwr <- CI.b0[1]
  b0.bs.upr <- CI.b0[2]
  b1.bs.lwr <- CI.b1[1]
  b1.bs.upr <- CI.b1[2]
  
  #Binding a record of bootstrapped estimates for beta0 to the result dataframe rdf 
  newrecord = data.frame(method="bootstrap", beta="beta0", est.value=b0.bs, SE=se.b0.bs, CI.lower=b0.bs.lwr, CI.upper=b0.bs.upr)
  rdf <- rbind(rdf, newrecord)
  #Binding a record of bootstrapped estimates for beta1 to the result dataframe rdf 
  newrecord = data.frame(method="bootstrap", beta="beta1", est.value=b1.bs, SE=se.b1.bs, CI.lower=b1.bs.lwr, CI.upper=b1.bs.upr)
  rdf <- rbind(rdf, newrecord)
  
  #Rounding off values
  rdf[, c(-1, -2)] <- round(rdf[, c(-1, -2)], 3)
  return(rdf)
}

#Testing the function
betaCoeffAnalysis(d, "HomeRange_km2~Body_mass_female_mean")
betaCoeffAnalysis(d, "MaxLongevity_m~Brain_Size_Species_Mean")
```

# Graph Plots {.tabset .tabset-fade .tabset-pills}
**Graph each beta value from the linear model and its corresponding mean value, lower CI and upper CI from a bootstrap as a function of number of bootstraps from 10 to 200 by 10s. HINT: the beta value from the linear model will be the same for all bootstraps and the mean beta value may not differ that much!**

## Preparing data
```{r}
#Preparing data frame for the final plots
plotDf <- data.frame(row.names = c("num.bootstraps", "lm.b0", "lm.b1", "bs.b0", "bs.b1", "bs.b0.lwr", "bs.b0.upr", "bs.b1.lwr", "bs.b1.upr"))

for (n in seq(from=10, to=200, by=10)) {
  df <- betaCoeffAnalysis(d, "MaxLongevity_m~Brain_Size_Species_Mean", conf.level = 0.95, n=n)
  
  newrecord <- data.frame(num.bootstraps=n, lm.b0=df$est.value[1], lm.b1=df$est.value[2], bs.b0=df$est.value[3], bs.b1=df$est.value[4], bs.b0.lwr=df$CI.lower[3], bs.b0.upr=df$CI.upper[3], bs.b1.lwr=df$CI.lower[4], bs.b1.upr=df$CI.upper[4])
  plotDf <- rbind(plotDf, newrecord)
}
#Dataframe comparing bootstrapped and lm() modeled beta coefficients and their CI intervals
plotDf
```

## lm() based values VS No. of Samples 
``` {r}
#Plot of lm() based values VS No. of Samples 
par(mfrow=c(1,2))
plot(y=plotDf$lm.b0, x=plotDf$num.bootstraps, ylab="Beta0 (lm)", xlab="no. of samples", main="Beta0 (lm) vs. No. of Samples")
xspline(x=plotDf$num.bootstraps, y=plotDf$lm.b0, shape=-0.4, lwd=2)

plot(y=plotDf$lm.b1, x=plotDf$num.bootstraps, ylab="Beta1 (lm)", xlab="no. of samples", main="Beta1 (lm) vs. No. of Samples")
xspline(x=plotDf$num.bootstraps, y=plotDf$lm.b1, shape=-0.4, lwd=2)
```

##Bootstrapped values VS No. of Samples 
```{r}
#Beta Coefficient Esimates
par(mfrow=c(1,2))
plot(y=plotDf$bs.b0, x=plotDf$num.bootstraps, ylab="Beta0", xlab="no. of samples", main="Beta0 vs. No. of Samples", ylim=c(230,260), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b0, shape=-0.4, lwd=2)

plot(y=plotDf$bs.b1, x=plotDf$num.bootstraps, ylab="Beta1", xlab="no. of samples", main="Beta1 vs. No. of Samples", ylim=c(1,3), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b1, shape=-0.4, lwd=2)

#Confidence Interval Esimates
par(mfrow=c(1,2))
plot(y=plotDf$bs.b0.lwr, x=plotDf$num.bootstraps, ylab="Beta0 CI (lower)", xlab="no. of samples", main="Beta0 CI (lower) vs. No. of Samples", ylim=c(150,250), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b0.lwr, shape=-0.4, lwd=2)

plot(y=plotDf$bs.b0.upr, x=plotDf$num.bootstraps, ylab="Beta0 CI (upper)", xlab="no. of samples", main="Beta0 CI (upper) vs. No. of Samples",ylim=c(150,250), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b0.upr, shape=-0.4, lwd=2)

par(mfrow=c(1,2))
plot(y=plotDf$bs.b1.lwr, x=plotDf$num.bootstraps, ylab="Beta1 CI (lower)", xlab="no. of samples", main="Beta1 CI (lower) vs. No. of Samples",ylim=c(0,2), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b1.lwr, shape=-0.4, lwd=2)

plot(y=plotDf$bs.b1.upr, x=plotDf$num.bootstraps, ylab="Beta1 CI (upper)", xlab="no. of samples", main="Beta1 CI (upper) vs. No. of Samples",ylim=c(0,2), xlim = c(5,200))
xspline(x=plotDf$num.bootstraps, y=plotDf$bs.b1.upr, shape=-0.4, lwd=2)
```